lifespan_eda

### 2024-11-15 eda/lifespan_eda.Rmd
adding plots to show site-level sex imbalances
clean up supplemental figs

```{r}
#just get first scan
cn.df_first <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice_head(n=1) %>%
  mutate(site_full = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="&")))) %>%
  ungroup() %>%
  group_by(site_full) %>%
  filter(n() > 1) #keep sites with at least 2 ppl
```

@@

```{r}
sfig_full_sample <- ggplot(cn.df_first) +
  geom_bar(aes( y=..count../tapply(..count.., ..x.. ,sum)[..x..], x=interaction(site, study, sep="&"), fill=sex)) +
  scale_y_continuous(labels = scales::percent) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(x = ggh4x::guide_axis_nested(delim="&")) +
  theme_linedraw() +
  theme(
    axis.ticks.x = element_blank(),
    axis.text.x = element_blank(),
    ggh4x.axis.nesttext.x = element_text(angle = 45, vjust = .9, hjust=1, size=5),
    ggh4x.axis.nestline = element_line(linetype = 1),
    legend.position="bottom"
  ) +
  geom_hline(yintercept=.5, color="lightgray", linetype="dashed") +
  labs(x="Data Collection Site", y="Sample Composition", fill="Sex")
ggsave(paste0(base_path, "/figs/sfig_full_sample.jpeg"), sfig_full_sample, width=15, height=7, units="in")
```

@@

p6 <- cn.df_first %>% 
  filter(study == "ADHD200") %>% 
  plot_by_sex() + + labs(title="ADHD200")

sfig_samples <- ggarrange(p1, p2, p3, p4, p5, p6, 
                          ncol=2, nrow=3, 
                          common.legend = TRUE, legend="right",
                          align="v")
ggsave(paste0(base_path, "/figs/sfig_subsamples.jpeg"), sfig_samples, width=15, height=10, units="in")
```

>>>>

```{r}
#just get first scan
cn.df_first <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice_head(n=1) %>%
  mutate(site_full = gsub(" ", "_", ifelse(is.na(site), study, paste(study, site, sep="&"))),
         age_yrs = (age_days-280)/365.25) %>%
  ungroup() %>%
  group_by(site_full) %>%
  filter(n() > 1) #keep sites with at least 2 ppl
```

@@

```{r}
sfig_full_sample <- ggplot(cn.df_first) +
  geom_bar(aes(x=..count../tapply(..count.., ..y.. ,sum)[..y..], y=interaction(site, study, sep="&"), fill=sex)) +
  scale_x_continuous(labels = scales::percent) +
  #theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1)) +
  guides(y = ggh4x::guide_axis_nested(delim="&")) +
  theme_linedraw() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    ggh4x.axis.nesttext.y = element_text(size=6),
    ggh4x.axis.nestline = element_line(linetype = 1),
    legend.position="right"
  ) +
  geom_vline(xintercept=.5, color="lightgray", linetype="dashed") +
  labs(y="Data Collection Site", x="Sample Composition", fill="Sex")
ggsave(paste0(base_path, "/figs/sfig_full_sample.jpeg"), sfig_full_sample, width=6, height=14, units="in")
```

@@

p6 <- cn.df_first %>% 
  filter(study == "ADHD200") %>% 
  plot_by_sex() + labs(title="ADHD200")

sfig_samples <- ggarrange(p1, p2, p3, p4, p5, p6, 
                          ncol=2, nrow=3, 
                          common.legend = TRUE, legend="right",
                          align="v",
                          labels = "AUTO")
ggsave(paste0(base_path, "/figs/sfig_subsamples.jpeg"), sfig_samples, 
       width=15, height=10, units="in")
```

look at age variability

```{r}
sfig_full_sample_age <- ggplot(cn.df_first) +
  geom_boxplot(aes(x=age_yrs, y=interaction(site, study, sep="&"))) +
  guides(y = ggh4x::guide_axis_nested(delim="&")) +
  theme_linedraw() +
  theme(
    axis.ticks.y = element_blank(),
    axis.text.y = element_blank(),
    ggh4x.axis.nesttext.y = element_text(size=6),
    ggh4x.axis.nestline = element_line(linetype = 1),
    axis.title.y=element_blank()
  ) +
  labs(x="Age (Years Post-Birth)")
ggsave(paste0(base_path, "/figs/sfig_full_sample_age.jpeg"), sfig_full_sample_age, width=6, height=14, units="in")

sitewise_full_plot <- ggarrange(sfig_full_sample, sfig_full_sample_age,
          ncol=2, nrow=1, align = "hv",
          common.legend = TRUE, legend = "bottom",
          labels="AUTO")
ggsave(paste0(base_path, "/figs/sitewise_full_plot.jpeg"), sitewise_full_plot, 
       width=15, height=15, units="in")
```

```{r}
plot_by_age <- function(df) {
  ggplot(df, aes(y = age_yrs, x=site)) +
    geom_boxplot() +
    theme_linedraw() +
    theme(legend.position = "none",
          axis.text.x = element_text(angle = 45, hjust = 1)) +
    xlab("Site") +
    ylab("Age")
}

p1.age <- cn.df_first %>% 
  filter(study == "ABCD") %>% 
  plot_by_age() + labs(title="ABCD")

p2.age <- cn.df_first %>% 
  filter(study == "UKB") %>% 
  plot_by_age() + labs(title="UK Biobank")

#bad balanced
p3.age <- cn.df_first %>% 
  filter(study == "EMBARC") %>% 
  plot_by_age() + labs(title="EMBARC")

p4.age <- cn.df_first %>% 
  filter(study == "ADNI") %>% 
  plot_by_age() + labs(title="ADNI")

p5.age <- cn.df_first %>% 
  filter(study == "abide1") %>% 
  plot_by_age() + labs(title="ABIDE 1")

p6.age <- cn.df_first %>% 
  filter(study == "ADHD200") %>% 
  plot_by_age() + labs(title="ADHD200")

sfig_samples.age <- ggarrange(p1.age, p2.age, p3.age, p4.age, p5.age, p6.age, 
                          ncol=2, nrow=3,
                          align="v",
                          labels = "AUTO")
ggsave(paste0(base_path, "/figs/sfig_subsamples_age.jpeg"), sfig_samples.age, 
       width=15, height=10, units="in")
```
